---
title: "Project proposal"
author: "zoomcoders"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction

*For our project, we want to investigate how the development of a country can impact the health of its citizens. For the purposes of this project, our research question will be "To what extent does economic development lead to better quality of life?"

Through our project, we aim to see whether more developped countries have better quality of life. While this would be expected, we will look to see if there are any instances where development has a negative correlation with certain elements of quality of life.

To analyse this, we will be using multiple datasets. As our dataset for development, we will use the Human Development Index by country, as that is a good indication of development in a country. For the other factors, we will look at datasets such as infant mortality by country, current health expenditure (% of GDP) by country, births attended by skilled health personnel by country, healthy life expectancy at birth (HALE) by country, maternity mortality ratio by country, leading causes of death by country, suicide rates by country, and population living below income poverty line, national poverty line (%) by country. 

Throughout this project, we aim to test the hypothesis that "Higher development leads to better quality of life".

Our datasets will be coming from:
- http://hdr.undp.org/en/indicators/137506
- http://hdr.undp.org/en/indicators/137506
- http://hdr.undp.org/en/indicators/137506
- https://www.who.int/data/gho/data/indicators/indicator-details/GHO/gho-ghe-hale-healthy-life-expectancy-at-birth
- https://www.who.int/data/gho/data/indicators/indicator-details/GHO/maternal-mortality-ratio-(per-100-000-live-births)
- https://www.who.int/data/gho/data/themes/mortality-and-global-health-estimates/ghe-leading-causes-of-death
- https://www.who.int/data/gho/data/themes/mental-health/suicide-rates
- https://www.who.int/data/gho/data/indicators/indicator-details/GHO/births-attended-by-skilled-health-personnel-(-)
Through our research question, we aim to see whether there is a strong positive relationship between the two variables as one would expect, or if it plays a smaller role in other scenarios.*


## 2. Data
```{r}

infant_mortality_rate <- read_csv("/cloud/project/data/InfantMortalityRate.csv")

human_development_index <- read_csv("/cloud/project/data/HumanDevelopmentIndex.csv")

healthy_life_expectancy_at_birth <- read_csv("/cloud/project/data/Healthy_life_expectancy_at_birth.csv")

maternal_mortality_ratio <- 
  read_csv("/cloud/project/data/maternal_mortality_ratio_per_100000_live_births.csv")
#Doesn't have values for 2019

births_attended_by_skilled_health_personnel <- read_csv("/cloud/project/data/births_attended_by_skilled_health_personnel.csv")
#Not fully up to date

suicide_rates_crude_10_year_age_groups <- read_csv("/cloud/project/data/suicide_rates_crude_10_year_age_groups.csv")

suicide_rates_age_standardized <- read_csv("/cloud/project/data/suicide_rates_age_standardized.csv")

glimpse(human_development_index)

glimpse(healthy_life_expectancy_at_birth)


# Data cleanup

infant_mortality_clean <- infant_mortality_rate %>%
  filter(Year == 2019 & Gender == "Total") %>%
  select(Country, `Infant Mortality Rate`)
#Remove "countries" such as "Central east Asia" and other continent names


hdi_clean <- human_development_index %>%
  select(2, 61)

colnames(hdi_clean) <- c("country", "hdi")

hdi_clean <- hdi_clean %>%
  filter(!is.na(country) & !(country == "Country") & !(country == "Human Development") & !(country == "Regions"))


hale_clean <- healthy_life_expectancy_at_birth %>%
  filter(Indicator == "Healthy life expectancy (HALE) at birth (years)" & Period == 2019 & Dim1 == "Both sexes") %>%
  select(Location, Value)


suicide_rates_age_groups <- suicide_rates_crude_10_year_age_groups %>%
  filter(Dim1 == "Both sexes") %>%
  select(Location, Dim2, Value)

suicide_rates_clean <- suicide_rates_age_groups %>%
  group_by(Location) %>%
  summarise(Value = mean(Value))
```




## 3. Data analysis plan

We will first merge all the data frames into one useable frame where each country has a value for each variable we have chosen. On the x of all the data representations we will plot either HDI, %of population below poverty line and %GDP expenditure on healthcare, as they are representative to development and on the y we will plot the rest as they are indicative of quality of life. The comparrison groups we will use will be countries. 
