---
title: "Project proposal"
author: "zoomcoders"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction

For our project, we want to investigate how the development of a country can impact the health of its citizens. For the purposes of this project, our research question will be "To what extent does economic development lead to better quality of life?"

Through our project, we aim to see whether more developped countries have better quality of life. While this would be expected, we will look to see if there are any instances where development has a negative correlation with certain elements of quality of life.

To analyse this, we will be using multiple datasets. As our dataset for development, we will use the Human Development Index by country, as that is a good indication of development in a country. For the other factors, we will look at datasets such as infant mortality by country, current health expenditure (% of GDP) by country, births attended by skilled health personnel by country, healthy life expectancy at birth (HALE) by country, maternity mortality ratio by country, leading causes of death by country, suicide rates by country, and population living below income poverty line, national poverty line (%) by country. 

Throughout this project, we aim to test the hypothesis that "Higher development leads to better quality of life".

Our datasets will be coming from:
- http://hdr.undp.org/en/indicators/137506
- http://hdr.undp.org/en/indicators/137506
- http://hdr.undp.org/en/indicators/137506
- https://www.who.int/data/gho/data/indicators/indicator-details/GHO/gho-ghe-hale-healthy-life-expectancy-at-birth
- https://www.who.int/data/gho/data/indicators/indicator-details/GHO/maternal-mortality-ratio-(per-100-000-live-births)
- https://www.who.int/data/gho/data/themes/mortality-and-global-health-estimates/ghe-leading-causes-of-death
- https://www.who.int/data/gho/data/themes/mental-health/suicide-rates
- https://www.who.int/data/gho/data/indicators/indicator-details/GHO/births-attended-by-skilled-health-personnel-(-)


## 2. Data
```{r}
countries_quality_of_life_index <- read_csv("/cloud/project/data/countries_quality_of_life_index.18-10-2021.csv") 

infant_mortality_rate <- read_csv("/cloud/project/data/InfantMortalityRate.csv")
```

> glimpse(countries_quality_of_life_index)
Rows: 115
Columns: 3
$ Name               <chr> "Canada", "Japan", "Norway", "Ireland", "Hungary", …
$ NativeName         <chr> "Canada", "日本", "Kongeriket Norge", "Eire", "Magy…
$ QualityOfLifeIndex <dbl> 159.51190, 165.32691, 173.15817, 153.18219, 135.343…

> glimpse(infant_mortality_rate)
Rows: 7,625
Columns: 4
$ Country                 <chr> "Afghanistan", "Angola", "Albania", "Andorra",…
$ `Infant Mortality Rate` <dbl> 43.050731, 44.851045, 7.659442, 2.555451, 5.71…
$ Gender                  <chr> "Female", "Female", "Female", "Female", "Femal…
$ Year                    <dbl> 2019, 2019, 2019, 2019, 2019, 2019, 2019, 2019…



## 3. Data analysis plan

We will plot "Quality of Life Index" on the Y axis and "Infant Mortality Rate" on the X axis, not grouping the data, since we have only 1 entry for both variables for each country. To prepare the data for what we want to use we will select the Total Infant mortality rates for 2019 and unite them with the Quality of Life data, so that we can have a new modified data set that has a value for "Country" corresponding to "Infant Mortality Rate" and "Quality of Life Index". Then we will plot a Jitter or Point graph with this new data set. To confirm or deny the relation we will need to check to what extent the points seem to follow a certain pattern. The stronger a pattern seems to be, the more sure we can be that there is some sort of relation between our chosen variables.
