---
title: "Project proposal"
author: "zoomcoders"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

*For instructions on what each section should include, please see the [project page](https://idsed.digital/assessments/project/#proposal) on the course website. Remove this text when completing your proposal*.

## 1. Introduction

For our project, we want to investigate the correlation between infant mortality rate and overall quality of life. We will do so by joining two datasets from the kaggle database; one on infant mortality rate trends by country, and the other being the quality of life index trends by country over the past ten years. Both these datasets are .csv files. 

Each dataset contains values of their respective variable for each country for a given number of years. Given that the infant mortality rate dataset has missing data for a lot of countries for most of the years, we will filter the dataset so that the time frame for both datasets shows only one year (2019), as this was the year where the most data was collected for infant mortality rate.

For the purposes of this project, our research question will be "To what extent does infant mortality rate impact quality of life?"

Through our research question, we aim to see whether there is a strong positive relationship between the two variables as one would expect, or if it plays a smaller role in other scenarios.

## 2. Data
```{r}
countries_quality_of_life_index <- read_csv("/cloud/project/data/countries_quality_of_life_index.18-10-2021.csv") 

infant_mortality_rate <- read_csv("/cloud/project/data/InfantMortalityRate.csv")
```

> glimpse(countries_quality_of_life_index)
Rows: 115
Columns: 3
$ Name               <chr> "Canada", "Japan", "Norway", "Ireland", "Hungary", …
$ NativeName         <chr> "Canada", "日本", "Kongeriket Norge", "Eire", "Magy…
$ QualityOfLifeIndex <dbl> 159.51190, 165.32691, 173.15817, 153.18219, 135.343…

> glimpse(infant_mortality_rate)
Rows: 7,625
Columns: 4
$ Country                 <chr> "Afghanistan", "Angola", "Albania", "Andorra",…
$ `Infant Mortality Rate` <dbl> 43.050731, 44.851045, 7.659442, 2.555451, 5.71…
$ Gender                  <chr> "Female", "Female", "Female", "Female", "Femal…
$ Year                    <dbl> 2019, 2019, 2019, 2019, 2019, 2019, 2019, 2019…



## 3. Data analysis plan

We will plot "Quality of Life Index" on the Y axis and "Infant Mortality Rate" on the X axis, not grouping the data, since we have only 1 entry for both variables for each country. To prepare the data for what we want to use we will select the Total Infant mortality rates for 2019 and unite them with the Quality of Life data, so that we can have a new modified data set that has a value for "Country" corresponding to "Infant Mortality Rate" and "Quality of Life Index". Then we will plot a Jitter or Point graph with this new data set. To confirm or deny the relation we will need to check to what extent the points seem to follow a certain pattern. The stronger a pattern seems to be, the more sure we can be that there is some sort of relation between our chosen variables.
