---
title: "Economic Development Effect on Quality of Life"
subtitle: <font size='6'>By countries, in 2019 </font>
author: <font size='8'>Zoomcoders <br> Alex Milchev, Francisco Mendonca, Jiaqi Gao and Yingying Xiao</font>
institute: <font size='10'>University of Edinburgh</font>
date: <font size = '6'>`r Sys.Date()`</font>
output:
  revealjs::revealjs_presentation:
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(rmarkdown)
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(revealjs)
library(data.table)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
dev_and_health <- read_csv(here::here("data/dev_and_health.csv"))
```

## Goal: To investigate how the development of a country can impact the quality of life of its citizens {data-background-color="rgb(0, 34, 51)"}
- Hypothesis: "If we plot different measurements of quality of life against development by country, we expect to see a positive linear relationship because higher development leads to better quality of life"


## Data {data-background-color="rgb(0, 34, 51)"}
- Data used in this analysis has come from two sources:
  - [UN Development Programme Human Development Reports](http://hdr.undp.org/)
  - [World Health Organization Global Health Observatory](https://www.who.int/data/gho/gho-search)

## Exploratory Analysis: Fitting Linear regressions {data-background-color="rgb(0, 34, 51)"}
- For the first part of our analysis, we chose to fit linear regressions to different measurements against HDI 
  - By doing this, we can firstly test our hypothesis and see if all these measurements have a positive linear relationship with HDI


# Life Expectancy vs HDI {data-background-color="rgb(0, 34, 51)"}
```{r Life Expectancy vs HDI, echo = FALSE}
life_expectancy_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(healthy_life_expectancy_at_birth ~ human_development_index, data = dev_and_health)
tidy(life_expectancy_HDI_fit)
  
glance(life_expectancy_HDI_fit)$adj.r.squared
```
## Life Expectancy vs HDI (Plot) {data-background-color="rgb(0, 34, 51)"}
$$\hat{Life Expectancy}_{i}=37.56912+36.04045\times HDI_{i}$$
```{r Life Expectancy vs HDI plot, out.width="70%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = healthy_life_expectancy_at_birth)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "HDI",
    y = "Life Expectancy",
    title = "Life Expectancy vs HDI")
```


# Births attended by skilled health personnel (%) vs HDI {data-background-color="rgb(0, 34, 51)"}
```{r Births attended by skilled health personnel rate vs HDI, echo = FALSE}
births_attended_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(births_attended ~ human_development_index, data = dev_and_health)
tidy(births_attended_HDI_fit)

glance(births_attended_HDI_fit)$adj.r.squared
```

## Births attended by skilled health personnel (%) vs HDI (Plot) {data-background-color="rgb(0, 34, 51)"}
$$\hat{health personnel}_{i}=24.32723+90.22275\times HDI_{i}$$
```{r Births attended by skilled health personnel rate vs HDI (Plot), out.width="60%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = births_attended)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(
    x = "HDI",
    y = "Births attended by skilled health personnel (%)",
    title = "Births attended by skilled health personnel rate vs HDI")
```

# Infant Mortality Rate vs HDI {data-background-color="rgb(0, 34, 51)"}
```{r Infant Mortality Rate vs HDI, echo = FALSE}
IMR_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(infant_mortality_rate ~ human_development_index, data = dev_and_health)
tidy(IMR_HDI_fit)

glance(IMR_HDI_fit)$adj.r.squared
```

## Infant Mortality Rate vs HDI (Plot) {data-background-color="rgb(0, 34, 51)"}
$$\hat{Infant Mortality Rate}_{i}=102.668-114.240\times HDI_{i}$$

```{r Infant Mortality Rate vs HDI (Plot), out.width="70%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = infant_mortality_rate)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(
    x = "HDI",
    y = "Infant Mortality Rate",
    title = "Infant Mortality Rate vs HDI")
```

# Number Maternal Deaths vs HDI {data-background-color="rgb(0, 34, 51)"}
```{r Number Maternal Deaths vs HDI, echo = FALSE}
NMD_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(number_maternal_deaths ~ human_development_index, data = dev_and_health)
tidy(NMD_HDI_fit)

glance(NMD_HDI_fit)$adj.r.squared
```

## Number Maternal Deaths vs HDI (Plot) {data-background-color="rgb(0, 34, 51)"}
$$\hat{Number Maternal Deaths}_{i}=9144. {data-background-color="rgb(0, 34, 102)"}853-10693.198\times HDI_{i}$$

```{r Number Maternal Deaths vs HDI (Plot), out.width="70%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = number_maternal_deaths)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "HDI",
    y = "Number Maternal Deaths",
    title = "Number Maternal Deaths vs HDI")
```

# Maternal Mortality Rate Per 100000 vs HDI {data-background-color="rgb(0, 34, 51)"}
```{r Maternal Mortality Rate Per 100000 vs HDI, echo = FALSE}
MMR_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(maternal_mortality_rate_per_100000 ~ human_development_index, data = dev_and_health)
tidy(MMR_HDI_fit)

glance(MMR_HDI_fit)$adj.r.squared
```

## Maternal Mortality Rate Per 100000 vs HDI (Plot) {data-background-color="rgb(0, 34, 51)"}
$$\hat{Maternal Mortality Rate Per 100000}_{i}=1027.427-1221.303\times HDI_{i}$$

```{r Maternal Mortality Rate Per 100000 vs HDI (Plot), out.width="60%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = maternal_mortality_rate_per_100000)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "HDI",
    y = "Maternal Mortality Rate Per 100000",
    title = "Maternal Mortality Rate Per 100000 vs HDI")
```

# Suicide Rate Per 100000 vs HDI {data-background-color="rgb(0, 34, 51)"}
```{r Suicide Rate Per 100000 vs HDI, echo = FALSE}
suicide_rate_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(suicide_rate_per_100000 ~ human_development_index, data = dev_and_health)
tidy(suicide_rate_HDI_fit)

glance(suicide_rate_HDI_fit)$adj.r.squared
```


## Suicide Rate Per 100000 vs HDI (Plot) {data-background-color="rgb(0, 34, 51)"}
$$\hat{Suicide Rate Per 100000}_{i}=62.91512-59.10957\times HDI_{i}$$

```{r Suicide Rate Per 100000 vs HDI (Plot), out.width="70%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = suicide_rate_per_100000)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "HDI",
    y = "Suicide Rate Per 100000",
    title = "Suicide Rate Per 100000 vs HDI")
```

# Initial analysis {data-background-color="rgb(0, 34, 51)"}
- From the initial regression fits, we can see that our hypothesis was correct
  - We generally saw a linear relationship between our measurements and HDI, seeing positive linear relationships with measurements such as healthy life expectancy, births attended by skilled health personnel, and negative linear relationships with measurements such as suicide rate, number of maternal deaths, infant mortality rate, and maternal mortality rate
  - However, the strength of these relationships varied, as we saw that the r squared values of some of these relationships was as high as 0.78, and some as low as 0.06.
    - For the next part of our analysis we want to see how the r squared values of these measurements compare to each other.

# Adjusted R Squared of Regression Model of Variables vs HDI {data-background-color="rgb(0, 34, 51)"}

```{r r_squared_HDI}
r_squared_HDI <- data.table(
  healthy_life_expectancy_at_birth = glance(life_expectancy_HDI_fit)$adj.r.squared,
  births_attended = glance(births_attended_HDI_fit)$adj.r.squared,
  infant_mortality_rate = glance(IMR_HDI_fit)$adj.r.squared,
  number_maternal_deaths = glance(NMD_HDI_fit)$adj.r.squared,
  maternal_mortality_rate_per_100000 = glance(MMR_HDI_fit)$adj.r.squared,
  suicide_rate_per_100000 = glance(suicide_rate_HDI_fit)$adj.r.squared
  ) %>%
  pivot_longer(
    cols = c(
      healthy_life_expectancy_at_birth,
      births_attended,
      infant_mortality_rate,
      number_maternal_deaths,
      maternal_mortality_rate_per_100000,
      suicide_rate_per_100000
    ),
    names_to = "variables",
    values_to = "adj_r_squared"
  )
r_squared_HDI
```

## Adjusted R Squared of Regression Model of Variables vs HDI (Plot) {data-background-color="rgb(0, 34, 51)"}

```{r adjusted r squared vs variables}

ggplot(data = r_squared_HDI, aes(y = variables, x = adj_r_squared, fill = variables)) +
  geom_bar(stat='identity') +
  scale_fill_manual(values = c(
    "healthy_life_expectancy_at_birth" = "#1b9e77",
    "births_attended" = "#d95f02",
    "infant_mortality_rate" = "#7570b3",
    "number_maternal_deaths" = "#e7298a",
    "maternal_mortality_rate_per_100000" = "#66a61e",
    "suicide_rate_per_100000" = "#e6ab02"
  ))+
  scale_x_continuous(limits = c(0, 1)) + 
  labs(
    y = "Health & Quality of Life Measurmenets",
    x = "Adjusted R Squared",
    title = "Adjusted R Squared Values",
    subtitle = "by Health & Quality of Life Measurements",
    fill = "Measurements")
  
```

# Adjusted R Squared Interpretation {data-background-color="rgb(0, 34, 51)"}
- We can see that variables such as healthy life expectancy at birth and infant mortality rates had the highest adjusted r^2 values.
  - We expect more developed countries to have better medical facilities, therefore have longer healthy life expectancies (strong positive relationship), and have lower rates of infant mortality (strong negative relationship)
- However, we also see that certain measurements such as maternal deaths and suicide rate have relatively low r squared values.
  - For maternal deaths, this can be explained by the fact that for HDI values higher than 0.8, number of maternal deaths was almost zero, hence not fitting a linear model that well
  - For suicide rate, we can infer that although medical facilities and technology may be quite high, this does not necessarily relate to high levels of mental health.

# Conclusion {data-background-color="rgb(0, 34, 51)"}
- We can overall say that our hypothesis was backed up by our results
  - However, the spread was much higher than expected
    - Highest r squared value for any measurement was 0.78, indicating that there isn't such a straightforward linear relationship between a measurement and HDI
- In addition we saw that measurements such as suicide rate don't necessarily have a strong relationship with HDI
  - This could be caused by a multitude of other factors, that may be more present in countries that are more developed
    - Example: stress caused by work, school, etc.

# Sources {data-background-color="rgb(0, 34, 51)"}
- [HDI by country - 2019](http://hdr.undp.org/en/indicators/137506)
- [Population living below poverty line (%) by country - 2019](http://hdr.undp.org/en/indicators/39006)
- [Current Health Expenditure (% of GDP) by country - 2019](http://hdr.undp.org/en/indicators/18180)
- [Maternity Mortality Ratio by country - 2019](https://www.who.int/data/gho/data/indicators/indicator-details/GHO/maternal-mortality-ratio-(per-100-000-live-births))
- [Suicide Rates by country - 2019](https://www.who.int/data/gho/data/themes/mental-health/suicide-rates)
- [Births attended by skilled personnel (%) by country - 2019](https://www.who.int/data/gho/data/indicators/indicator-details/GHO/births-attended-by-skilled-health-personnel-(-))
- [Healthy Life Expectancy by country - 2019](https://www.who.int/data/gho/data/indicators/indicator-details/GHO/gho-ghe-hale-healthy-life-expectancy-at-birth)
