---
title: "Economic Development Effect on Quality of Life"
subtitle: <font size='6'>By countries, in 2019 </font>
author: <font size='8'>Zoomcoders <br> Alex Milchev, Francisco Mendonca, Jiaqi Gao and Yingying Xiao</font>
institute: <font size='10'>University of Edinburgh</font>
date: <font size = '6'>`r Sys.Date()`</font>
output:
  revealjs::revealjs_presentation:
    self_contained: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(rmarkdown)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(revealjs)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
dev_and_health <- read_csv(here::here("data/dev_and_health.csv"))
```

## A statement of the overall goal / research question {data-background-color="rgb(169, 169, 169)"}

# Section title

# Hello World

- Click the `Knit` button to compile your presentation

- Make sure to commit and push all resulting files to your GitHub repo

# Using reveal.js

It's possible to have a horizontal sliding transition as well as nest multiple slides to create a vertical stack.

When presenting, you use the left/right arrows to step through the top-level (horizontal) slides. When you arrive at a vertical stack you can optionally press the up/down arrows to view the vertical slides or skip past them by pressing the right arrow.

(click the down key now)

## reveal.js

- The presentation is created using the `reveal.js` package

- You can create a slide show broken up into sections by using the # and ## heading tags

- You can also create a new slide without a header using a horizontal rule (----)

----

- You can also add pauses between content on a slide using (". . .")

. . .

Like this.

...or using "::: fragment"

::: fragment

Like this.

:::

# Layouts

You can use plain text

- or bullet points

<div>
  <div style="float: left; width: 50%;">or text in two columns</div>
  <div style="float: left; width: 50%;">like this</div>
</div>

because you can combine HTML and Markdown syntax.

<p style="font-size: 8pt; position: absolute; bottom: -5em; width: 100%; text-align: center;">*You can also add footnotes</p>

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
  tidy(model)
  glance(model)
```

# Life Expectancy vs HDI
```{r Life Expectancy vs HDI, echo = FALSE}
life_expectancy_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(healthy_life_expectancy_at_birth ~ human_development_index, data = dev_and_health)
tidy(life_expectancy_HDI_fit)
  
glance(life_expectancy_HDI_fit)$adj.r.squared
```
## Life Expectancy vs HDI (Plot)
$$\hat{Life Expectancy}_{i}=37.56912+36.04045\times HDI_{i}$$
```{r Life Expectancy vs HDI plot, out.width="70%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = healthy_life_expectancy_at_birth)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "HDI",
    y = "Life Expectancy",
    title = "Life Expectancy vs HDI")
```


# Births attended by skilled health personnel (%) vs HDI
```{r Births attended by skilled health personnel rate vs HDI, echo = FALSE}
births_attended_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(births_attended ~ human_development_index, data = dev_and_health)
tidy(births_attended_HDI_fit)

glance(births_attended_HDI_fit)$adj.r.squared
```

## Births attended by skilled health personnel (%) vs HDI (Plot)
$$\hat{health personnel}_{i}=24.32723+90.22275\times HDI_{i}$$
```{r Births attended by skilled health personnel rate vs HDI (Plot), out.width="60%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = births_attended)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(
    x = "HDI",
    y = "Births attended by skilled health personnel (%)")
```

# Infant Mortality Rate vs HDI
```{r Infant Mortality Rate vs HDI, echo = FALSE}
IMR_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(infant_mortality_rate ~ human_development_index, data = dev_and_health)
tidy(IMR_HDI_fit)

glance(IMR_HDI_fit)$adj.r.squared
```

## Infant Mortality Rate vs HDI (Plot)
$$\hat{Infant Mortality Rate}_{i}=102.668-114.240\times HDI_{i}$$

```{r Infant Mortality Rate vs HDI (Plot), out.width="70%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = infant_mortality_rate)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(
    x = "HDI",
    y = "Infant Mortality Rate")
```

# Number Maternal Deaths vs HDI
```{r Number Maternal Deaths vs HDI, echo = FALSE}
NMD_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(number_maternal_deaths ~ human_development_index, data = dev_and_health)
tidy(NMD_HDI_fit)

glance(NMD_HDI_fit)$adj.r.squared
```

## Number Maternal Deaths vs HDI (Plot)
$$\hat{Number Maternal Deaths}_{i}=9144.853-10693.198\times HDI_{i}$$

```{r Number Maternal Deaths vs HDI (Plot), out.width="70%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = number_maternal_deaths)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "HDI",
    y = "Number Maternal Deaths")
```

# Maternal Mortality Rate Per 100000 vs HDI
```{r Maternal Mortality Rate Per 100000 vs HDI, echo = FALSE}
MMR_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(maternal_mortality_rate_per_100000 ~ human_development_index, data = dev_and_health)
tidy(MMR_HDI_fit)

glance(MMR_HDI_fit)$adj.r.squared
```

## Maternal Mortality Rate Per 100000 vs HDI (Plot)
$$\hat{Maternal Mortality Rate Per 100000}_{i}=1027.427-1221.303\times HDI_{i}$$

```{r Maternal Mortality Rate Per 100000 vs HDI (Plot), out.width="60%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = maternal_mortality_rate_per_100000)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "HDI",
    y = "Maternal Mortality Rate Per 100000")
```

# Suicide Rate Per 100000 vs HDI
```{r Suicide Rate Per 100000 vs HDI, echo = FALSE}
suicide_rate_HDI_fit <- linear_reg() %>%
	set_engine("lm") %>%
	fit(suicide_rate_per_100000 ~ human_development_index, data = dev_and_health)
tidy(suicide_rate_HDI_fit)

glance(suicide_rate_HDI_fit)$adj.r.squared
```


## Suicide Rate Per 100000 vs HDI (Plot)
$$\hat{Suicide Rate Per 100000}_{i}=62.91512-59.10957\times HDI_{i}$$

```{r Suicide Rate Per 100000 vs HDI (Plot), out.width="70%", fig.width=5, echo = FALSE}
ggplot(data = dev_and_health, aes(x = human_development_index, y = suicide_rate_per_100000)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "HDI",
    y = "Suicide Rate Per 100000")
```

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

## Plot and text

::: columns
:::: column
* Some text
* goes here
::::

:::: column
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
::::
:::

## Different widths

::: columns
:::: {.column width="40%"}
* Some text
* goes here
::::

:::: {.column width="60%"}
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
::::
:::

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

## {data-background-image="https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80"}

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs (\$...\$), e.g. $\alpha+\beta$. You can use the display style with double dollar signs:

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

# Feeling adventurous?

You can learn more about the options available at the following pages:

- https://github.com/rstudio/revealjs

- https://bookdown.org/yihui/rmarkdown/revealjs.html

- http://www.jenniferbradham.org/post/reveal_js_markdown/

- https://revealjs.com/